---
title: "UMMC Cleaning Workflow"
author: "C Murphy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## UMMC Transfusion Data Cleaning

This workflow is designed to identify blood products from a database that uses a combination of ISBT terminology and short strings pulled from the ISBT product descriptions. This depends on having access to a full list of ISBT blood product codes, which are copyrighted but should be accessbile from your local IT or billing department.

```{r}
## Read in a csv of ISBT codes that have been previously matched to product IDs (see the ISBTCodeCleaner RMD).
ISBTCodes <- read.csv("ISBT_codes_productID.csv")

## Assign product types via ISBT codes and unique short descriptive strings. ^ (caret) is a regex operator that indicates the pattern to be matched is at the beginning of the line.
redCellCodes <- ISBTCodes %>% filter(ProductID == "RBC")
redCellCodes <- paste(c(as.character(redCellCodes$Ecode), "RC/", "RC1/", "RC2/", "W/Aph/RC/", "W/RC/", "Deg/RC/", "Deg/Aph/RC/", "Aph/RC/"), collapse = "|^")
head(redCellCodes)

plasmaCodes <- ISBTCodes %>% filter(ProductID == "FFP")
plasmaCodes <- paste(c(as.character(plasmaCodes$Ecode), "T/P", "T/Aph", "T/C/P", "D6720", "D6730"), collapse = "|^")
head(plasmaCodes)

plateletCodes <- ISBTCodes %>% filter(ProductID == "PLT")
plateletCodes <- paste(c(as.character(plateletCodes$Ecode), "A/P", "W/A/P"), collapse = "|^")
head(plateletCodes)

cryoCodes <- ISBTCodes %>% filter(ProductID == "CRYO")
cryoCodes <- paste(c(as.character(cryoCodes$Ecode)), collapse = "|^")
head(cryoCodes)

```


```{r echo = FALSE}
## Import data
transfusionData <- read.csv("YOUR DATA HERE.csv")

## Convert dates from factors into date objects via stripping the time and converting to a POSIXCT
transfusionData <- transfusionData %>% mutate(Issued_DateTime_POSIXct = as.POSIXct(strptime(EVENT_DT_TM, format = "%m/%d/%Y %H:%M"))) %>% mutate(dayOfWeek = weekdays(Issued_DateTime_POSIXct)) %>% mutate(Issued_DateOnly_POSIXct = as.POSIXct(strptime(format(Issued_DateTime_POSIXct, "%m/%d/%Y"), "%m/%d/%Y")))

## Our database has multiple institutions. This code selects the institution of interest. 
##It turns out that a lot of the older data has FACILITY as blank when it's mostly UMMC.
transfusionData <- transfusionData %>% filter(FACILITY == "" | FACILITY == "Univ MD Med Sys")

## ProductID label applied
transfusionData <- transfusionData %>% mutate(ProductID = case_when(str_detect(PRODUCT_TYPE_DISPLAY, redCellCodes) ~ "RBC", str_detect(PRODUCT_TYPE_DISPLAY, plasmaCodes) ~ "FFP", str_detect(PRODUCT_TYPE_DISPLAY, plateletCodes) ~ "PLT", str_detect(PRODUCT_TYPE_DISPLAY, cryoCodes) ~ "CRYO", TRUE ~ "Missing"))

## Taking only the columns we want for analysis
transfusionData <- transfusionData %>% select(PATIENT_NAME, FIN, MRN, PATIENT_ABO_CD, PATIENT_RH_CD, NURSE_UNIT, DISPENSED_TO_LOCATION, PRODUCT_NBR, ProductID, PRODUCT_TYPE_DISPLAY, Product_ABORH, TRANSFUSED_VOL, Issued_DateTime_POSIXct, Issued_DateOnly_POSIXct, dayOfWeek, Dispensed_Physician_NAME)

## Ouptut the file!
write.csv(transfusionData, "YOUR OUTPUT NAME HERE.csv")
head(transfusionData)

```
